---
title: "PS270 Final Project"
author: "Allison Breitzmann"
description: "My Final Project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to campaign financing, specifically where the money comes from and its relation to electability. Either in terms of ideology of those donations or the size of the donation themselves.

Research Question: How does the size of donation impact the electability of candidates?

Does size of donations impact a candidates electability? In this study, I plan to analyze the extent in which the percent of campaign donations from individuals under $200 influences the final election results. I hypothesize that a higher percent of small donations will result in a larger percent of the vote share. Those with a greater amount of smaller donations are better able to mobilize and attract a larger amount of individuals donating, whereas those with large donations focus more so on PACs and lobbying groups which can not cast a ballot when the time comes. My sample includes the House and Senate nominees from the 2024 Election. My explanatory variable is the percentage of total donations coming from donations of 200 dollars or less. The variable will be split into ranges of 25%. My outcome variable will be the average percentage of vote share. If a positive correlation is found between percentage of small donations and percent of vote share, this would support my hypothesis. However, if larger donations result in higher voteshare, than my hypothesis would be incorrect.

```{r include = FALSE}
library(tidyverse)
library(broom)
```


```{r include = FALSE}
Donations <- read_csv("data/Large vs. Small Donations, Current General Election Candidates, 2024 Cycle.csv")
head(Donations)
Vote_Share <- read_csv("data/Nominees-2024.csv") |>
  select(Nominee, Race, Type, Party, Votes, Share, Total)
head(Vote_Share)
```




```{r}
Donations |>
  mutate(
    clean_name = str_replace(Name, " \\(.\\)", ""),
    clean_name = str_replace(clean_name, ".* ", ""),
    .before = Name
  ) #|>
 # write_csv("donation.csv")
```

```{r}
Vote_Share |>
  mutate(clean.name = str_replace(Nominee, ".* ", ""),
         .before = Nominee) #|>
  #write_csv("vote.csv")
```

```{r include = FALSE}
clean_Donations <- read_csv("donation.csv")
clean_Vote_Share <- read_csv("vote.csv")
```

```{r}
Vote_Link <- Vote_Share |>
  inner_join(clean_Vote_Share)
Donation_Link <- Donations |>
  inner_join(clean_Donations)
Data <- Donation_Link |>
  inner_join(Vote_Link)
```
```{r}
Data <- Data |>
  mutate(Percent_Small_Donors = str_replace(`Percent from Small Donors*`,"%",""), .before = `total money raised`) |>
  mutate(Percent_Small_Donors = as.integer(Percent_Small_Donors)) |>
  mutate(`Total from Small Donors` = str_replace(`Total from Small Donors`, "$",""))
```


```{r include = FALSE}
Data <- Data |>
  mutate(Total_Small = str_replace(`Total from Small Donors`, "\\$", ""), .after = `Total from Small Donors` ) |>
  mutate(Total_Small = as.integer(Total_Small)) |>
  mutate(`Total Amount Raised` = case_when(Total_Small < 10000 ~ "1. Less Than $10,000 Raised",
                                   Total_Small > 10000 & Total_Small < 100000 ~ "2. Between $10,000 and $100,000 Raised",
                                   Total_Small > 50000 & Total_Small < 10000000 ~ "3. Between $100,000 and $1,000,000 Raised",
                                   Total_Small > 1000000 ~ "4. Over $1,000,000 Raised")) |>
  mutate(Total_Small_tenthousand = Total_Small/10000,
         Share_perc = Share*100)
```

```{r include = FALSE}
Data <- Data |>
  mutate(PercentSmall = case_when(Percent_Small_Donors < 10 ~ "1. Less than 10% from Small Donors",
                                  Percent_Small_Donors >= 10 & Percent_Small_Donors < 25 ~ "2. Between 10% and 25% from Small Donors",
                                  Percent_Small_Donors >= 25 & Percent_Small_Donors < 50 ~ "3. Between 25% and 50% from Small Donors",
                                  Percent_Small_Donors >= 50 & Percent_Small_Donors < 75 ~ "4. Between 50% and 75% from Small Donors",
                                  Percent_Small_Donors >= 75 ~ "Over 75% from Small Donors"))
Data
```



```{r}
ggplot(Data, aes(x = Percent_Small_Donors,
                 y = Share_perc)) +
  geom_point(aes(color = `Total Amount Raised`)) + 
  geom_smooth(method = "lm") +
  scale_x_sqrt() +
  labs(title = "Relationship between Vote Share and Percentage of Small Donations",
       x = "Percentage of Individual Donations less than $200",
       y = "Percentage of Vote Share")
```

```{r}
Regression <- lm((Share*100) ~ Percent_Small_Donors + Total_Small_tenthousand, data = Data)
Regression_sum <- glance(Regression)
Regression |>
  tidy() |>
  mutate(adj.r.squared = Regression_sum$adj.r.squared) |>
   mutate(Coefficients = case_when(term == "(Intercept)" ~ "Intercept",
                          term == "Percent_Small_Donors" ~ "Percentage of Individual Donations less than $200",
                          term == "Total_Small_tenthousand" ~ "Total Amount Raised (ten thousands)")) |>
   select(Coefficients, estimate, std.error, adj.r.squared) |>
  knitr::kable()
  
```

```{r}
Regression
```


When a Candidate receives 0% of the Vote Share, it is predicted the Candidate will receive 52.37% of the Vote Share. When the total amount raised remains constant, it is estimated that for every additional percentage point of donations of less than $200, the candidate will lose 0.35 percentage points of the Vote Share. However, with an adjusted R^2 value of 0.075, only 7.5% of the variance is explained by the model, meaning the linear regression has a poor model fit and very little correlation between Vote Share and Percentage of Individual Donations under 200 dollars